---
title: "Performance Prediction"
author: "Albert Buchard"
date: "15 September 2016"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(data.table)
setwd("~/Google Drive/dev/R/pipelineProject/r-pipeline/examples")
pathToScriptFolder = "../R" 

## Load processed data for grouped analysis and correlation data
source(file.path(pathToScriptFolder, "modeling.R"))
source(file.path(pathToScriptFolder, "utilities.R"))

```

# Prediction of performance evolution in learning a determinist trajectory
4*4 grid
Minimum period 7 transition, max 16
Supposing you start in the main cycle
Memory and learning strategy getting better with the levels 
```{r}

numberOfSubjects = 200
performanceDT = NULL
for (level in seq(1,30,1)) {
  memory = logb(level, 30)*0.65+0.15
  cyclePeriod = sample(seq(7,16),numberOfSubjects, T)
  performance = sapply(seq(0,25,1), function(x) {
    if (x<7) {
      return(ntrue(sample(seq(0,1,0.01),NROW(cyclePeriod),T)<0.25)/NROW(cyclePeriod))
    }
    sampleOneCycleVector = sample(seq(0,1,0.01),ntrue(cyclePeriod<=x),T)
    sampleRandomChoiceVector = sample(seq(0,1,0.01),ntrue(sampleOneCycleVector>memory)+ntrue(cyclePeriod>x),T)
    
    return((ntrue(sampleOneCycleVector<=memory)+ntrue(sampleRandomChoiceVector<0.25))/NROW(cyclePeriod))
  })
  performanceDT = rbind(performanceDT, data.table(level=level,transition=seq(0,25,1),performance=performance))
  # sum of bool choice good / number of subjects
}
ggplot(performanceDT[level<13,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)
ggplot(performanceDT[level>=13,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)

# Using non uniform cycle size distribution (approximated with gaussian (but should not...)
numberOfSubjects = 200
simulateLearning = function (numberOfSubjects = 200, condition =1) {
  performanceDT = NULL
  for (level in seq(1,30,1)) {
    #memory = logb(level, 30)*0.65+0.15
    memory = rnorm(numberOfSubjects, logb(level, 30)*0.5+0.25, sd=0.2*(1-(level/300)))
    memory[memory>1] = 1
    memory[memory<0] = 0
    
    cyclePeriod = rnorm (numberOfSubjects, mean = 7, sd = 4) #sample(seq(7,16),numberOfSubjects, T)
    cyclePeriod[cyclePeriod<7] = 7 + abs(7-cyclePeriod[cyclePeriod<7])
    cyclePeriod = floor(cyclePeriod)
    
    
    if (condition == 1) {
      performance = sapply(seq(0,25,1), function(x) {
        if (x<7) {
          return(ntrue(sample(seq(0,1,0.01),NROW(cyclePeriod),T)<0.25)/NROW(cyclePeriod))
        }
        sampleOneCycleVector = sample(seq(0,1,0.01),ntrue(cyclePeriod<=x),T)
        sampleRandomChoiceVector = sample(seq(0,1,0.01),ntrue(sampleOneCycleVector>memory[cyclePeriod<=x])+ntrue(cyclePeriod>x),T)
        
        return((ntrue(sampleOneCycleVector<=memory[cyclePeriod<=x])+ntrue(sampleRandomChoiceVector<0.25))/NROW(cyclePeriod))
      })
    }
    
    if (condition == 2) {
      # same rule for each color
      numberOfLearnedRules = floor(rnorm(numberOfSubjects, logb(level, 30)*3, sd=1-(level/300)))
      numberOfLearnedRules[numberOfLearnedRules>4] = 4
      numberOfLearnedRules[numberOfLearnedRules<0] = 0
      print(numberOfLearnedRules)
      
      ## there's a way to introduce bias due choose same rule for the random color if the random rule is not learned 
      ## subject will try to find a common rule for this color if he did not learn
      learnedRandomRule = sample(seq(0,1,0.05), numberOfSubjects, T) < logb(level, 30)*rnorm(numberOfSubjects, 0.8, sd=1-(level/300))
      
      performance = sapply(seq(0,25,1), function(x) {
        colorType = sample(seq(1,5,1), numberOfSubjects,T)
        isRuleLearned = numberOfLearnedRules>=colorType
        
        
        sampleOneCycleVector = sample(seq(0,1,0.01),ntrue(cyclePeriod[isRuleLearned==F]<=x),T)
        memoryForNotLearned = memory[isRuleLearned==F]
        sampleRandomChoiceVector = sample(seq(0,1,0.01),
                                        ntrue(sampleOneCycleVector>memoryForNotLearned[cyclePeriod[isRuleLearned==F]<=x])+
                                          ntrue(cyclePeriod[isRuleLearned==F]>x),T)
        # print("------")
        # print(NROW(isRuleLearned))
        # print(NROW(memoryForNotLearned))
        #  print("*")
        # print(NROW(sampleOneCycleVector))
        # print(NROW(sampleRandomChoiceVector))
        # print(ntrue(isRuleLearned))
        
        return((ntrue(sampleOneCycleVector<=memoryForNotLearned[cyclePeriod[isRuleLearned==F]<=x])+
                  ntrue(sampleRandomChoiceVector<0.25)+
                  ntrue(isRuleLearned))/NROW(cyclePeriod))
      })
      
      
    }
    
    performanceDT = rbind(performanceDT, data.table(level=level,transition=seq(0,25,1),performance=performance))
    # sum of bool choice good / number of subjects
  }
  return(performanceDT)
}

performanceDT200 = simulateLearning(numberOfSubjects)

ggplot(performanceDT200[level<13,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)
ggplot(performanceDT200[level>=13&level<=24,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)

numberOfSubjects = 1
performanceDTOne = simulateLearning(numberOfSubjects)
performanceDTOne[,choice:= "Correct"]
performanceDTOne[performance==0, choice := "Incorrect"]
performanceDTOne$choice = factor(performanceDTOne$choice, levels = c("Incorrect","Correct"))
ggplot(performanceDTOne[level<13,], aes(x=transition, y=choice)) + geom_point() + facet_wrap(~level, ncol=4)
ggplot(performanceDTOne[level%in%c(1,15,30),], aes(x=transition, y=choice)) + geom_point() + facet_wrap(~level, ncol=4)

performanceDT200Cond2 = simulateLearning(numberOfSubjects, 2)

ggplot(performanceDT200Cond2[level<13,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)
ggplot(performanceDT200Cond2[level>=13&level<=24,], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)
ggplot(performanceDT200Cond2[level%in%c(1,15,30),], aes(x=transition, y=performance)) + geom_point() + facet_wrap(~level, ncol=4)

for (level in seq(1,30,1)) {
  if (level >20) {
    lambda = 1
  } else {
    if (level<5) {
      lambda = level/10
    } else {
      lambda = level/20 + 0.25
    }
    
  }
  performance = data.table(x=seq(0,25,1), y=c(rep(0.25,7),sapply(seq(0,18,1), function(x) { return(sigmoid(x, c(0.6+level/100, lambda, 13-(level*0.5))) * 0.75 + 0.25) } )))
  toPlot = ggplot(performance, aes(x=x,y=y)) + geom_point()
  print(toPlot)
}
```

# Prediction of performance evolution in general rule condition
4*4 grid
Minimum period 7 transition
Supposing you start in the main cycle
Memory and learning strategy getting better with the levels 
```{r}
for (level in seq(1,30,1)) {
  if (level >20) {
    lambda = 5
  } else {
    if (level<5) {
      lambda = level/5
    } else {
      lambda = 5*level/20
    }
    
  }
  performance = data.table(x=seq(0,25,1), y=sapply(seq(0,25,1), function(x) { return(sigmoid(x, c(1, lambda, 12-(level*0.2))) * 0.75 + 0.25) } ))
  toPlot = ggplot(performance, aes(x=x,y=y)) + geom_point()
  print(toPlot)
}
```
